Diversity dynamics in Nymphalidae butterflies: Effect of phylogenetic uncertainty on diversification rate shift estimates

### 2014-12-19 response_to_reviewers.tex

\section{Response to reviewers}\label{response-to-reviewers}

\section{Reviewer 1}\label{reviewer-1}

\textbf{1. Reviewer's Comment:}

 Page 3, last sentence of the first paragraph in the Introduction is a fragment; consider joining it to
preceding sentence with a colon (:).

@@

\textbf{1. Response:}

\begin{quote}
\color{blue}
We have followed the reviewer's advise.
\end{quote}

\textbf{2. Reviewer's Comment:}

Page 3, last sentence of Introduction: as this is the first mention of BiSSE inbe spelled out (even though it is spelled out in the Abstract as well).

@@

\textbf{2. Response:}

\begin{quote}
\color{blue}
We have followed the reviewer's advise.
\end{quote}

\textbf{3. Reviewer's Comment:}

 Page 3, first paragraph of Data section in Materials & Methods, final sentence: replace “correct for”
with “account for”.

@@

\textbf{3. Response:}

\begin{quote}
\color{blue}
We have followed the reviewer's advise.
\end{quote}

\textbf{4. Reviewer's Comment:}

Page 3, second paragraph of Data section in Materials & Methods, first sentence: the two web
resources (Tree of Life Web Project and Global Butterfly Names project) should be listed in the
References section and treated as references in the manuscript (i.e. referred to by number).

@@

\textbf{4. Response:}

\begin{quote}
\color{blue}
We have moved the web links to the references as advised by the reviewer.
\end{quote}

@@

\textbf{5. Reviewer's Comment:}

Page 4, Analyses of Diversification section, second paragraph: The BAMM analyses are given less
than cursory treatment, and no real comparison is discussed in the Discussion (different
diversification rate shifts are identified in the BAMM analyses than in the MultiMEDUSA
analyses). At this point, either a flushed-out discussion of the potential reasons for the differences
should be included, or the BAMM analyses should just be removed from the manuscript (the latter
would likely be best).

@@

\textbf{5. Response:}

\begin{quote}
\color{blue}
We have removed the BAMM analysis from the manuscript.
\end{quote}

@@

\begin{quote}
\color{blue}
``Because most of Nymphalidae butterflies are restricted to use one
plant family as hostplant (232 genera use only one plant family versus
176 that use more than one; based on our data in supp. mat. 15), the
\ldots{}''
\end{quote}

\textbf{6. Reviewer's Comment:}

Page 4, Detecting diversification shifts... section: A brief (one to two sentences) description of how
models become iteratively more complex in MEDUSA analyses is warranted here. Something
about how one additional clade-specific diversification rate parameter is added at each step.

@@

\textbf{6. Response:}

\begin{quote}
\color{blue}
We have rewritten the paragraph and included additional sentences as requested
by the reviewer.
\end{quote}

\textbf{7. Reviewer's Comment:}
Page 4, Estimation of trait-dependent... section, first sentence: replace “binary state speciation and extinction” with “BiSSE model”.

\begin{quote}
\color{blue}
We have followed the reviewer's advise.
\end{quote}

\textbf{8. Reviewer's Comment:}

 Page 4, Estimation of trait-dependent... section, third sentence: it is not clear why the reference to MuSSE is necessary here, as the authors do not appear to use it. If this is the case, the sentence could be omitted.

 \textbf{8. Response:}

\begin{quote}
\color{blue}
We removed that sentence.
\end{quote}

\textbf{9. Reviewer's Comment:}
Page 5, Estimation of trait-dependent... section, sentence beginning “Other hostplant shifts were not tested...”: is this accurate? At the end of the results section, the very last sentence describes a test for the effect of Apocynaceae feeding on diversification rates...

\textbf{9. Response:}

\begin{quote}
\color{blue}
We removed that sentence.
\end{quote}

### 2014-12-19 MS_revised.tex
comment 8
comment 9

As the MEDUSA and MultiMEDUSA approaches estimated an increase in net
diversification in the clade Ithomiini, we tested whether this pattern
can be explained by an increase in the birth-rate due to hostplant use
and performed analyses using the BiSSE model \cite{maddison2007} 
as implemented in the R package
\texttt{diversitree} \cite{fitzjohn2012}. The Markov Chain Monte Carlo algorithm
was run for 10000 generations discarding the first 7500 as burnin. The
Multiple State Speciation and Extinction approach (MuSSE) \cite{fitzjohn2012} is
designed to examine the joint effects of two or more traits on
speciation. Because most of Nymphalidae butterflies are restricted to
use one plant family as hostplant (232 genera use only one plant family
versus 176 that use more than one; based on our data in supp. mat. 15),
the character states can be coded as presence/absence, for which the
BiSSE analysis is better suited. BiSSE was designed to test whether a
binary character state has had any effect on increased net
diversification rate for a clade \cite{maddison2007}. We used our compiled data of
hostplant use to produce a binary dataset for the character ``feeding on
the plant family Solanaceae'' with two states (absence = 0; presence =
1) (supp. mat. 17). The family Solanaceae include the hostplants of the
diverse Ithomiini butterflies and closest relatives \cite{willmott2006}. Other
hostplant shifts were not tested for effect on net diversification rates
due to the low support for diversification shifts found in the MEDUSA
analyses. We analyzed the data using BiSSE employing the Markov Chain
Monte Carlo algorithm on the maximum clade credibility tree, taking into
account missing taxa by using the parameter ``sampling factor''
(\texttt{sampling.f}) in \texttt{diversitree}. We also used constrained
analyses forcing no effect of hostplant use on diversification and used
likelihood ratio tests to find out whether the hypothesis of effect on
diversification has a significantly better likelihood than the null
hypothesis (no effect). The analyses were run across a sample of 250
trees from the selected 1000 random trees from the posterior
distribution. The records of \emph{Vanessa} and \emph{Hypanartia}
feeding on Solanaceae \cite{beccaloni2008, scott1986} might be incorrect as it is unlikely
that these species can be feeding on this plant family. We tested
whether coding these two genera as not feeding on Solanaceae affected
our results. We expected that coding these two genera as either absence
or presence would not have a significant effect on our results. It has
been shown that BiSSE performs poorly under certain conditions \cite{davis2013}.
However, our data has adequate number of taxa under analysis (more than
300 tips), adequate speciation bias (between 1.5x and 2.0x), character
state bias (around 8x) and extinction bias (around 4x) for the analysis
of Solanaceae hostplants. Thus, BiSSE is expected to produce robust
results \cite{davis2013}.

>>>>

As the MEDUSA and MultiMEDUSA approaches estimated an increase in net
diversification in the clade Ithomiini, we tested whether this pattern
can be explained by an increase in the birth-rate due to hostplant use
and performed analyses using the BiSSE model \cite{maddison2007} 
as implemented in the R package
\texttt{diversitree} \cite{fitzjohn2012}. The Markov Chain Monte Carlo algorithm
was run for 10000 generations discarding the first 7500 as burnin.
Because most of Nymphalidae butterflies are restricted to
use one plant family as hostplant (232 genera use only one plant family
versus 176 that use more than one; based on our data in supp. mat. 15),
the character states can be coded as presence/absence, for which the
BiSSE analysis is better suited. BiSSE was designed to test whether a
binary character state has had any effect on increased net
diversification rate for a clade \cite{maddison2007}. We used our compiled data of
hostplant use to produce a binary dataset for the character ``feeding on
the plant family Solanaceae'' with two states (absence = 0; presence =
1) (supp. mat. 17). The family Solanaceae include the hostplants of the
diverse Ithomiini butterflies and closest relatives \cite{willmott2006}. 
We analyzed the data using BiSSE employing the Markov Chain
Monte Carlo algorithm on the maximum clade credibility tree, taking into
account missing taxa by using the parameter ``sampling factor''
(\texttt{sampling.f}) in \texttt{diversitree}. We also used constrained
analyses forcing no effect of hostplant use on diversification and used
likelihood ratio tests to find out whether the hypothesis of effect on
diversification has a significantly better likelihood than the null
hypothesis (no effect). The analyses were run across a sample of 250
trees from the selected 1000 random trees from the posterior
distribution. The records of \emph{Vanessa} and \emph{Hypanartia}
feeding on Solanaceae \cite{beccaloni2008, scott1986} might be incorrect as it is unlikely
that these species can be feeding on this plant family. We tested
whether coding these two genera as not feeding on Solanaceae affected
our results. We expected that coding these two genera as either absence
or presence would not have a significant effect on our results. It has
been shown that BiSSE performs poorly under certain conditions \cite{davis2013}.
However, our data has adequate number of taxa under analysis (more than
300 tips), adequate speciation bias (between 1.5x and 2.0x), character
state bias (around 8x) and extinction bias (around 4x) for the analysis
of Solanaceae hostplants. Thus, BiSSE is expected to produce robust
results \cite{davis2013}.
